---
title: Simulate dose titration in the model
author: Metrum Research Group, LLC
output:
  html_document:
    css: styles/styles.css
    toc: true
    number_sections: true
    footer: metrumrg &copy 2017
---
```{r,echo = FALSE}
knitr::opts_chunk$set(comment='.', message = FALSE, warning = FALSE)
```


# The model

- One-compartment with extravascular absorption
    - CL = 1, VC = 20, KA = 1.2
    - No IIV
- Direct link Emax model
    - E0 = 100, Emax = 60, EC50 = 25
    - 30% IIV on EC50 and 10% IIV on E0

# The titration scheme

- Everyone gets 100 mg q12h for a week
- After that, start monitoring the trough response and adjust doses 
according to a dosing protocol
    - `118 < RESP < 122` keep the dose the same
    - `122 <  RESP < 150` decrease by 20%
    - `RESP > 150` decrease by 40%
    - `105 < RESP < 118` increase by 30%
    - `RESP < 105` double the dose 
- Implement the titration protocol for six weeks and 
summarize baseline and final response at week 10

# To investigate

- Summarize the distribution of final responses relative 
to baseline responses both with and without the dose titration
scheme in play

## A twist to try out

- Instead of adjusting the dose based on the trough response,
try adjusting the dose based on the mean of the last 
3 daily response measurements

# Setup 

```{r}
library(mrgsolve)
library(dplyr)
library(ggplot2)
library(tidyr)
```


We have written more code


```{r}
code <- '
$SET req = ""

$GLOBAL
double F1  = 1;

double adjust(double resp, double F1) {
  if(resp > 118 && resp < 122) return F1;
  if(resp > 122 && resp < 150) return F1 * 0.8;
  if(resp > 150) return F1 * 0.6;
  if(resp > 105 && resp < 118) return F1 * 1.3;
  if(resp < 105) return F1 * 2;
  return F1;
}

bool tta(double time, double evid, double titr) {
  if(titr !=1 || evid !=0 ) return false;
  return  (time/168 == floor(time/168)) && time > 7*24 && time <= 7*7*24;
}

$PARAM 
TVCL = 1, V = 20, KA = 1.2, 
TVEC50 = 20, TVE0 = 100
TITR = 1

$PKMODEL cmt = "GUT CENT", depot = TRUE

$OMEGA 0 0.09 100

$MAIN 

if(NEWIND <=1 ) F1 = 1;
double CL = TVCL*exp(ETA(1));
double EC50 = TVEC50*exp(ETA(2));
double E0 = TVE0 + ETA(3);

F_GUT = F1;

$TABLE
capture CP = CENT/V;
capture RESP = E0 + 60*CP/(EC50+CP);

if(tta(TIME,EVID,TITR)) F1 = adjust(RESP,F1); 

capture F1out = F1;

'
```


The simulation duration
```{r}
dur <- 10*7*24
```


```{r}
mod <- 
  mcode("foo", code) %>% 
  update(end = dur) %>% 
  ev(amt=100, ii=12, addl=2*10*7)

mod <- param(mod, TITR = 1)
```


## Simulate 50 individuals

```{r}
set.seed(102214)
out <- mrgsim(mod, nid = 50, delta = 2, obsonly = TRUE)
plot(out)
```


## Summarize the simulations

- Just get the rows at baseline and at the end
- Derive a `variable` column for plotting

```{r}
sum <- 
  filter(out, time %in% c(0,dur)) %>% 
  group_by(ID) %>% mutate(variable = c("Baseline", "Week 10"))
```

```{r}
ggplot(data=sum) + 
  geom_density(aes(x=RESP, fill=variable), alpha = 0.6)
```

```{r}
set.seed(102214)
out <- 
  mod %>% param(TITR = 0) %>% 
  mrgsim(nid = 50, delta = 1, obsonly = TRUE)

plot(out)
```


```{r}
sum <- 
  filter(out, time %in% c(0,dur)) %>% 
  group_by(ID) %>% mutate(variable = c("Baseline", "Week 10"))
```

```{r}
ggplot(data=sum) + 
  geom_density(aes(x=RESP, fill=variable), alpha = 0.6)
```


# Another approach

- Keep the code in file `adaptive.h` in the `project` directory
- in `$GLOBAL`, create an object for tracking history
- A `reset` method gets called at the start of the problem
and at the start of every individual
- At the right time, we can save some information
and get at that information as well as other historical 
information

```{r}
code <- '
$INCLUDE adaptive.h

$GLOBAL
hx obj;

$PREAMBLE 
obj.reset();

$CMT A

$ODE dxdt_A = -0.5*A;

$MAIN 
if(NEWIND <= 1) obj.reset();

$TABLE
obj.save(A);

capture previous = obj.previous();
capture current = obj.current();

'
```

```{r}
mod <- mcode("foo", code, project = 'model', soloc = '.')

mrgsim(init(mod, A = 100))
```


