---
title: "Annotated model specification"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    toc: true
    css: docs/src/styles/styles.css
---

```{r,echo = FALSE}
knitr::opts_chunk$set(comment='.', warning = FALSE, message = FALSE)
```


# Introduction

Annotated model specification is an alternate format for selected 
blocks in the model specification file. 

The "annotated" block allows the user to enter an extended 
set of information about selected model elements and also 
allows the user to create a summary of the model contents 
based on that extended data set.

## What can be annotated

- Parameters
- Compartments and initial condtions
- Output variables
- Problem summmary


# Annotated parameter block

Let's start with an annotated parameter block to introduce the 
different formatting considerations.

The __standard__ paramater block looks like
```{c, eval = FALSE}
$PARAM CL = 1, VC = 2, KA = 3
```

The format is: `name = value`.  It would be nice if 
we could also include a brief description of the 
parameter as well as the unit.

The equivalent parameter block in __annotated__ format would be
```{c, eval = FALSE}
$PARAM @annotated
CL : 1 : Clearance (L/hr)
VC : 2 : Volume of distribution (L)
KA : 3 : Absorption rate constant (1/hr)
```


__Of note__:

1. We still use the `$PARAM` block identifier to specify the parameters
1. We also use the `@annotated` directive to identify that the block 
is in annotated format.
1. For each parameter, specify the:
    * parameter name (`CL`)
    * colon
    * parameter value (`1`)
    * colon
    * parameter decription (`Clearance`)
    * parameter unit (`(L/hr)`)
1. In the __standard__ specification, we could put multiple parameters 
on the same line; in the __annotated__ specification, each parameter 
needs its own line
  
## Unit specification 

The parameter unit is found inside the __last__ set of parentheses
found on the specification line. 

So you could do this
```{c,eval = FALSE}
$PARAM @annotated
CL : 1 : Clearance (healty volunteers) (L/hr)
```

If you don't want a unit either don't write anything 
```{c,eval = FALSE}
$PARAM @annotated
F1 : 1 : Bioavailability fraction
```

or put in a "no-value" value; `mrgsolve` uses a `.` to hold the value

```{c,eval = FALSE}
$PARAM @annotated
F1 : 1 : Bioavailability fraction (tablet) (.)
```

# Compartments

Recall that we can specify compartments with either

* `$CMT name1 name2` just lists the compartment names, assumes values are 0
* `$INIT name1 = value1, name2 = value2` give the name and value
an `$INIT` block.  

To annotate `$CMT`

```{c, eval = FALSE}
$CMT @annotated
DEPOT   : SC dosing compartment (nmol)
CENTRAL : Central compartment (nmol)
```

__Of note__:

1. All of the details that we discussed for `$PARAM` hold for `$CMT`
1. __Except__ we omit the value of the compartment (consistent with 
what we do in the __standard__ specification

To annotate `$INIT`

```{c,eval = FALSE}
$CMT @annotated
DEPOT   : 0 : SC dosing compartment (nmol)
CENTRAL : 0 : Central compartment (nmol)
```

__Of note__:

1. The same as `$CMT` except we __do__ include __value__
1. All elements are analgous to what we discussed for `$PARAM`


# Annotate output variables

Recall that we can derive numeric values in the model code and mark 
those values to be included in the simulated __output__

For example
```{c,eval = FALSE}
$TABLE
double CP = CENT/VC;

$CAPTURE CP
```

We can make an annotated version of the `$CAPTURE` block like this

```{c, eval=FALSE}
$CAPTURE @annotated
CP : Concentration (ng/mL)
```

Like `$CMT`, `$CAPTURE` only takes the names of the items 
to be included in the output.  So we construct the __annotated__
specification blocks with the `name : description (unit)` format.


# Problem specification

There is a block in the `mrgsolve` specification where we can include 
details about the model run.  Frequently, we include a short description
of the run. 

```{c, eval = FALSE}
$PROB 
Final PK model run
```


In the annotated approach, we can include information in markdown 
format

```{c,eval = FALSE}
$PROB
- Summary: PK model run
- Covariates: Weight, CLCR
- Random effects:
    - CL: exponential
    - VC: exponential
- Error model: combined
- Derived from: run 1002
```




# Extract model details 

When we do follow an annotated model specification format, we can 
extract that information from the model object

```{r}
code <- '
$PARAM @annotated
A : 2 : first parameter (ng)
B : 5 : second parameter (kg)

$CMT @annotated
DEPOT : Dosing compartment
CENTRAL : Central compartment

$TABLE
double C = 10;
double D = 20;

$CAPTURE @annotated
C : Equals 10 (kg/m2)
D : Equals 20 (dL)
'

library(mrgsolve)
mod <- mcode("foo", code)
```


Use `mrgsolve:::details` to extract the information

```{r}
mrgsolve:::details(mod)
```

The `mrgsolve:::details` function returns a data frame that tells you the 

* block
* name
* description
* value
* unit

where applicable for each annotated item


# Render the document
```{r,eval}
mod <- mread("irm1", modlib())

mrgsolve:::details(mod)
```


Use `mrgsolve:::render` to create a document out of the model.

```{r, eval = FALSE, message = FALSE}
mrgsolve:::render(mod)
```


















