---
title: "Pediatric dose evaluation"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document: 
    toc: true
    css: docs/src/styles/styles.css
    footer: metrumrg &copy 2017
---

```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE)
```

# Packages

- Loading some packages
- Sourcing in some code in `functions.R`

```{r}
a <- config::get()
source(a$functions)
library(dplyr)
library(ggplot2)
library(mrgsolve)
library(tidyr)
```

# Publication

- Title: *Population pharmacokinetics of azithromycin and chloroquine in healthy adults and paediatric malaria subjects following oral administration of fixed-dose azithromycin and chloroquine combination tablets*
- Authors: Qinying Zhao, Thomas G Tensfeldt, Richa Chandra, Diane R Mould
- Reference: Malaria Journal 2014, 13:36

## Conclusion
"Weight was predictive of exposure for both AZ and CQ. Assuming equivalent dosing (mg/kg), AZ and CQ exposure in children would be expected to be lower than that in adults, suggesting that children may require a higher dose (mg/kg) than adults to achieve the same AZ and CQ exposure."


## Our objective
- Explore this conclusion through simulation
- If we used the typical adult mg/kg dose in a pediatric population, 
is the AUC really that different?  How much different?

# Simulated ages and weight

```{r}
data <- readRDS(a$az_datafile) 

head(data)
```

Derive a grouping variable based on `AGE`

- 1 = `0.5 <= AGE < 5`
- 2 = `5 <= AGE <= 12`
- 3 = `AGE >= 18`

```{r}
data <- mutate(data, 
               group = case_when( AGE >= 0.5 & AGE < 5 ~ 1, 
                                  AGE >= 5 & AGE <= 12 ~ 2,
                                  AGE >=18 ~ 3))

data <- filter(data, !is.na(group))
```


# Azithrymycin dosing

Modify this covariate data frame so that we have 

- oral, bolus doses
- daily dosing over 3 days

```{r}
data <- mutate(data, cmt=1, ii=24, addl=2, time=0, evid=1) %>% as_data_frame
```

* Sample 2000 subjects in each of the age groups

```{r}
data <- group_by(data, group) %>% sample_n(2000) %>% ungroup
head(data)
```

Assign doses

- For the pediatric cohorts, 300 mg if weight > 20 kg and 100 mg otherwise
- For adults, 500 mg
- doses are daily x 3 days

```{r}
data <- mutate(data, dose = if_else(WT > 20, 300, 100))
data <- mutate(data, dose = if_else(group==3, 500, dose), dosekg = dose/WT)
```

The doses were assigned into `dose`; we need to copy that into
`amt` as well

```{r}
data <- mutate(data, amt = dose)
head(data)
```


# Azithromycin model
```{r}
mod <- mread(a$az_model, a$az_project)

see(mod)
```


Update the model object to cover 3 days of dosing
```{r}
mod <- update(mod, end=72, delta=0.2)
```

## Simulate all age groups

```{r}
sims <- mrgsim(mod, data = data, obsonly = TRUE, Req = "CP")
```


Some post-processing to calculate the `AUC` on `DAY` 3
```{r}
sims <- mutate(sims, DAY = 1+floor(time/24)) %>% filter(DAY==3)

auc <- AUC(sims, time="time", dv="CP") %>% left_join(data)
```

Now, dose normalize the AUC (and switch up the units)
```{r}
auc <- mutate(auc, ratio  = AUC/dosekg/1000)

head(auc)
```

Mark the adult group for the plot
```{r}
auc <- mutate(auc, pop  = if_else(group==3, "adult", "pediatric"))
```



## Plot
```{r}
ggplot(data=auc, aes(x=pop, y=ratio)) + geom_boxplot()
```



# Your turn
 
Find the typical per/kg dose in the adult population and 
apply that dose to the 6 - 59 month population
 
```{r}
auc %>% group_by(group) %>% summarise(mean = mean(dosekg))
```

Let's take 7 mg/kg and simulate out 6 to 59 month old kids next to adults
What is the distribution of the adult/pediatric auc ratio

```{r}
data4 <- filter(data, group != 2) %>% mutate(amt = WT*7)
```

```{r}

sims <- mrgsim(mod, data = data4, obsonly = TRUE, Req = "CP")

sims <- mutate(sims, DAY = 1 + floor(time/24)) %>% filter(DAY==3)

auc <- AUC(sims, time="time", dv="CP") 

auc <- 
  left_join(auc, data4) %>% 
  group_by(group) %>%
  mutate(unit = 1:n()) %>%
  ungroup

auc <- 
  dplyr::select(auc, unit, AUC, group ) %>% 
  spread(group, AUC, sep = "")

auc <- mutate(auc, ratio = group1/group3)
```

Summarize

```{r}
ggplot(data=auc, aes(x=ratio)) + geom_density() + xlim(0,5)

summarize(auc, 
          mean = mean(ratio), 
          lower = quantile(ratio, 0.025), 
          upper = quantile(ratio, 0.975))
```

```{r}
mean(auc$ratio < 1)
```

