---
title: "Categorical R / Rcpp"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document: 
    css: docs/src/styles/styles.css
    toc: true
    number_sections: true
---

```{r,echo=FALSE}
source("libPaths.R")
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE)
```

# Packages

```{r}
library(dplyr)
library(mrgsolve)
library(readr)
library(magrittr)
library(ggplot2)
library(parallel)
```



# A simple categorical covariate simulation

- Using `$PLUGIN Rcpp` to get at that `R` namespace
- Allows us to get all of the `dpqr` functions

```{r}
code <- '
$PLUGIN Rcpp
$MAIN
int i = R::rbinom(1, 0.5);
$CAPTURE i
'
```

- This takes a little longer to compile compared to models without the Rcpp `$PLUGIN`

```{r}
mod <- mcode_cache("categorical", code)
```


```{r}
out <- mrgsim(mod, end = 10000)
out
mean(out$i)
```


# Now, we want multinomial simulation

- See `?rmultinom`
- Distribute `size` objects into `K` boxes


# First, an `R`-based solution

- Here, we will call the `R` function `rmultinom` from our model code
- Performance hit, but we can mitigate that if we do it right

## First, let's see how to use an `R` function in the model

```{r}
code_print <- '
$PLUGIN Rcpp

$GLOBAL
Rcpp::Function prnt("print");

$MAIN
if(TIME==3) {
  prnt("The time is 3");
  prnt(TIME);
}
'
```

- And run this
```{r}
mod <- mcode_cache("print", code_print)

mrgsim(mod, end = 5)
```


## Back to the `R`-based solution


```{r}
code3 <- '
$PARAM n = 100

$PLUGIN Rcpp mrgx

$GLOBAL
Rcpp::Function rmulti("rmultinom");
Rcpp::NumericVector probs;
Rcpp::NumericVector ans;

$ENV
probs <- c(0.25, 0.25, 0.5)

$PREAMBLE
probs = mrgx::get<Rcpp::NumericVector>("probs", self);

$MAIN
ans = rmulti(1, n, probs);

$TABLE
capture ans1 = ans(0,0);
capture ans2 = ans(1,0);
capture ans3 = ans(2,0);
'
```

- This time we really call back to `R` and use the `R` version of the function
    - The signature is the same as the one we use in `R`
    - The return object (matrix) is also the same as what et get in `R`
- This time, we also invoke `$PLUGIN mrgx` along with `$ENV`
    - We can specify the `probs` vector in `$ENV`
        - You write plain old `R` code in `$ENV`
    - We can __get__ that `probs` vector through `mrgx::get` 
        - This is a templated function ... we have to tell the function 
        the __type__ of object we are expecting to get back
        - Notice that the call is `mrgx::get<Rcpp::NumericVector>get(...)`, 
        we are assigning that to `probs`, which is also a `Rcpp::NumericVector`
        - Mysterious `self` object; just think of that as the location where we hide
        the address of what is in `$ENV`


- Run this

```{r}
mod <- mcode_cache("categorical3", code3)
out <- mrgsim(mod, param = list(n = 100))
out
```

```{r}
out <- mrgsim(mod, param = list(n = 10000))
out
```


Work a little with the contents of `$ENV`

```{r}
env_ls(mod)
```

```{r}
mod@envir
```

```{r}
env_get(mod)
```


- Update our `probs` vector and resimulate

```{r}
mod <- env_update(mod, probs = c(0.1, 0.7, 0.2))
mrgsim(mod, param = list(n = 10000))
```


- You can stash any `R` object in there and go in and get it later
- Best to stick to `numeric`, `matrix` and `vector`
    - `data.frame` and `list` are hard to work with 

## A more-realistic implementation

- Here, we pre-simulate every individual in `$PREAMBLE`

```{r}
code3b <- '
$PARAM n = 100

$PLUGIN Rcpp mrgx

$GLOBAL
Rcpp::Function rmulti("rmultinom");
Rcpp::NumericVector probs;
Rcpp::NumericMatrix ans;

$ENV
probs <- c(0.25, 0.25, 0.5)

$PREAMBLE
probs = mrgx::get<Rcpp::NumericVector>("probs", self);

ans = rmulti(self.nid, 1, probs);

$MAIN
capture ans1 = ans(0,self.idn);
capture ans2 = ans(1,self.idn);
capture ans3 = ans(2,self.idn);
'
```

- Simulate to see what happens in 1000 IDs

```{r}
mod <- mcode_cache("categorical3b", code3b)
out <- mrgsim(mod, end = -1, nid = 1000)
out
```

- Summarize to confirm
```{r}
mean(out$ans1)
mean(out$ans2)
mean(out$ans3)
```


### Create a function to return the group number

- Name `get_group`
- Returns `int`
- Arguments: the matrix of sims and the ID number
    - Pass matrix by reference


```{r, eval = FALSE}
code3c <- '
$PARAM n = 100

$PLUGIN Rcpp mrgx

$GLOBAL
Rcpp::Function rmulti("rmultinom");
Rcpp::NumericVector probs;
Rcpp::NumericMatrix ans;

int get_group(Rcpp::NumericMatrix& ans, int idn) {
  if(ans(0,idn)==1) return 1;
  if(ans(1,idn)==1) return 2;
  if(ans(2,idn)==1) return 3;
  return 0;
}

$ENV
probs <- c(0.25, 0.25, 0.5)

$PREAMBLE
probs = mrgx::get<Rcpp::NumericVector>("probs", self);
ans = rmulti(self.nid, 1, probs);


$MAIN
capture group = get_group(ans,self.idn);
capture ans1 = ans(0,self.idn);
capture ans2 = ans(1,self.idn);
capture ans3 = ans(2,self.idn);
'
```


```{r, eval = FALSE}
mod <- mcode_cache("categorical3c", code3c)
out <- mrgsim(mod, end = -1, nid = 1000)
out
```
```{r, eval = FALSE}
filter(out, group ==2)
```


# Here is a fully-`C++` implementation

- This is harder to do, but performance will be better
- Note that `rmultinom` is different from most of the other `dqpr` functions
    - I had to go find this solution on `SO`

```{r}
code2 <- '
$PARAM n = 100

$PLUGIN Rcpp

$GLOBAL
Rcpp::NumericVector prob(3);
Rcpp::IntegerVector ans(3);
int k;

$PREAMBLE
prob[0] = 0.25;
prob[1] = 0.25;
prob[2] = 0.5;
k = int(prob.size());

$MAIN
R::rmultinom(n, prob.begin(), k, ans.begin());

$TABLE
capture ans1 = ans[0];
capture ans2 = ans[1];
capture ans3 = ans[2];
'
```

## What's going on 

- Declare our data objects in `$GLOBAL`
    - Unlike `R`, when we want a vector to be numeric or integer, we have to explicitly 
    state that
    - Declare with the name(size) (e.g. `Rcpp::NumericVector foo(10)`)
- Initialize their values in `$PREAMBLE`
    - `prob` we set to `c(0.25, 0.25, 0.5)`
    - `k` we want to be the number of boxes
- In `$MAIN` we actually call the `rmultinom` function
    - This pattern is odd compared to the other functions, but it's what we have to do
    - The last argument is a pointer to the `ans` vector; the `answer` gets filled in there; 
    there is no return value to `rmultinom`
- In `$CAPTURE`, we just collect the `ans` values

```{r}
mod <- mcode_cache("categorical2", code2)
out <- mrgsim(mod, param = list(n = 1000))
out
```


