---
title: "Flexible designs"
output:
  html_document:
    css: docs/src/styles/styles.css
    toc: true
    number_sections: true
---

```{r,echo=FALSE,message=FALSE}
knitr::opts_chunk$set(comment='.', fig.align=TRUE,message=FALSE,warning=FALSE)
```


# Packages
```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
```


# Introduction

We are free to create data sets which (1) include
observation records in addition to dosing records and (2) 
there is no limit to the person-to-person heterogeneity 
in those dosing designs.  

However, it is not at all convenient to program this 
variability into the data set.  We would rather have 
a programmatic way to create a list of designs and 
assign different designs to different individuals 
or groups of individuals.  

## Objective

- Review `tgrid` objects, what they are and how are they used
- Introduce methods for creating heterogeneity in observation
schedule within a given simulation run

<hr><div class = spacer></div>

# tgrid objects

- `mrgsolve` tracks a simulation `start` and `end` time as well 
as a `delta` value specifying how often an observation will be made
between `start` and `end`
- There is also an additional, ad-hoc vector of simulation times (`add`)
- `tgrid` objects abstract that functionality and allow us to create separate
objects that describe a period of time using those same `start/end/delta/add` 
elements
- Recall also that we can work with `tgrid` objects through operations, the 
most common being `+` ... a shift in time

```{r}
mod <- mread_cache("popex", modlib())

day1 <- tgrid(0,24,2)
day5 <- day1 + 5*24

des <- c(day1,day5)

des
```

## Default behavior
Regardless how we use the `tgrid` object, every individual 
in the simulation is assigned the same grid of observation times

```{r}
mod %>% 
  ev(amt=100, ii=24, addl = 9) %>%
  mrgsim(tgrid = des, nid = 10) %>% 
  plot(DV~., type = 'b')
```

<hr><div class = spacer></div>

# Assign unique end times to each ID

- Create an `idata_set` including each `ID` in the run
  - column includes `end` ... this will be the simulation 
  end time 

```{r}
idata <- data_frame(ID = 1:100) 
idata <- mutate(idata, end = sample(c(1,2,3,4)*168, n(), replace = TRUE))
idata
```

Use `as_deslist` to create a list of designs ... 

- We look the column indicated by `descol` to understand how many 
different designs we need to make. 
- By taking `descol = "end"`, there will be 4 different designs 
by virtue of the fact that we've got 4 different end times

```{r}
des <- as_deslist(idata, descol = "end")
des
```

What whould have happened if `descol = "ID"`?


Now, simulate with that design list
```{r}
out <- 
  mod %>% 
  idata_set(idata) %>%
  ev(amt=100, ii=24, addl = 28) %>%
  design(des) %>% 
  mrgsim(carry.out = "end")
```

It is required that we

1. implement an `idata_set` ... that is where the information is coming from and 
1. identify the `idata_set` prior to identifying the `deslist`

```{r}
plot(out, DV~time|factor(end))
```

What if we wanted less intense samples in the 2, 3, 4 week designs
but not the 1 week design?

```{r}

```


<hr><div class = spacer></div>


