---
title: "Fit PK data in R"
author: ""
date: ""
output: 
  html_document:
    css: docs/src/styles/styles.css
    toc: true
    number_sections: true
---

```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE)
set.seed(11010)
```

# Packages
```{r}
library(minqa)
library(RcppDE)
library(ggplot2)
library(mrgsolve)
library(dplyr)
library(purrr)
```

<hr><div class = spacer></div>

# Fit simple, single-subject PK data `R`

Go to the `Theoph` data set for an example

- The `exTheoph` data set gets shipped with `mrgsolve`


```{r}
data(exTheoph)
```


We'll just pick off the first `ID` to fit
```{r}
data <- filter(exTheoph, ID==1)
obs  <- filter(data, evid==0)
dose <- filter(data, evid==1)
```

```{r}
qplot(data=data, time,conc, geom=c("line", "point"))
```



These are the observed data 

```{r}
yobs <- obs$conc %>% unlist %>% unname
yobs
```

I'm going to capture them as a just a vector of 
values for now.  


Like in the warmup vignette, we create a function 
that takes care of getting the prediction
and calculating the objective function (ols, here).

- `p` parameters
- `.mod` model object
- `.data` the estimation data set
- `.yobs` vector of observed data
- `pred` if `TRUE`, just return a prediction


```{r}
ols <- function(p, .mod, .data, .yobs, pred=FALSE) {
  
  p <- lapply(p, exp)
  
  .mod <- update(.mod, param=p)
  
  if(pred) {
    .data <- filter(.data, evid !=0)
    return(mrgsim(.mod,.data))
  }
  
  out <- mrgsim(.mod, data=.data, obsonly=TRUE)
  
  y.hat <- out$CP
  
  sum(c(y.hat-.yobs)^2)
}
```

The workflow is similar to the warmup vignette, but __some important differences__

1. We have the parameters on log scale here, so we need to untransform
1. We cannot use the parameters directly to get the prediction; we 
need to perform an `update` step to bring the parameter set being proposed 
by the optimizer into the model object
1. We use `mrgsim` to actually generate the prediction
    - What would happen if we simulated with `obsonly = FALSE`?
1. When we call `ols` with `pred = TRUE`, we can get the prediction


Pull a simple 1-cmt model from the library
```{r}
mod <- mread("pk1", modlib())
```


```{r}
see(mod)
```

Initial estimates; we'll keep the parameters on log-scale

```{r}
theta <- log(c(CL=1, V=1, KA=0.1))
```


Let's introduce a new optimizer from the `minqa` package
called `newuoa`.

```{r}
fit <- minqa::newuoa(theta, ols, .mod = mod, .data=data, .yobs = yobs)

est <- setNames(fit$par, names(theta))

exp(fit$par)
```



Get a  prediction at the __final__ estimate

```{r}
pred <- 
  mod %>% 
  update(end=25, delta=0.1) %>%
  ols(est,.,data,yobs,pred=TRUE) %>% as.data.frame
```


Get a prediction at the __initial__ estimate
```{r}
pred_theta <- 
  mod %>% 
  update(end=25, delta=0.1) %>%
  ols(theta,.,data,yobs,pred=TRUE) %>% 
  as_data_frame
```


Plot of observed data and data predicted at the 
initial and final estimates
```{r}
ggplot() + 
  geom_line(data=pred, aes(time,CP),lwd=1, col="firebrick") +
  geom_point(data=data, aes(time,conc), size=3, col="darkslateblue") +
  geom_line(data=pred_theta, aes(time,CP), lwd=1, lty=2,col="darkgreen")
```


# Speed up the simulation with `qsim`

```{r}
ols.q <- function(p,.mod,.e,.data,.yobs, pred=FALSE) {
  
  p <- lapply(p, exp)
  
  .mod <- update(.mod,param=p)
  
  if(pred) {
    .data <- filter(.data, evid!=0)
    return(mrgsim(.mod,.data))
  }
  
  out <- qsim(.mod, .e)
  
  y.hat <- out[,"CP"]
  
  sum(c(y.hat-.yobs)^2)
}
```


```{r}
datam <- recmatrix(as.ev(dose),times=obs$time)
yhat <- qsim(mod,datam)
```


## Benchmark 

```{r}
e1 <- expression(newuoa(theta,ols.q,.mod=mod,.e=datam,.data=data,.yobs=yobs))
e2 <- expression(newuoa(theta,ols,.mod=mod,.data=data,.yobs=yobs))

rbenchmark::benchmark(eval(e1), eval(e2),replications=5)[,1:4]
```

 
# The optimhelp package

This is in progress, but I almost always use it when working problems of any size. 


We can make a "set" of parameters  like this
```{r}
library(optimhelp)


cl <- log_par("CL", 1)
vc <- log_par("VC", 10)

p <- parset(cl,vc)

```



```{r}
p
```

So we have two parameters (`CL` and `VC`) with parameters (1 and 10) and they are both on log scale.


To start an estimation problem, call
```{r}
initials(p)
```

