--- 
title: "Pooled analysis of indomethacin PK data"
author: Metrum Research Group, LLC
date: ""
output: 
  html_document:
    css: docs/src/styles/styles.css
---

# Packages

```{r}
source("libPaths.R")

library(ggplot2)
library(dplyr)
library(mrgsolve)
```

# Load indomethacin data set
```{r}
data(Indometh)
```

- Take a look at what is there
```{r}
head(Indometh)

count(Indometh, Subject)
```

# Data assembly 

- Assume single 25 mg dose for all individuals
- Setting `cmt` to 0 for observations
- Setting `cmt` to 2 for doses ... will need a model 
where the second compartment is `central`

```{r}
obs <- as.data.frame(Indometh) %>% 
  mutate(evid = 0, cmt = 0, ID = as.numeric(Subject)) 

dose <- 
  distinct(obs, ID) %>% 
  mutate(amt = 25, time = 0, conc = NA, evid = 1, cmt = 2)
```

- Put it back together 

```{r}
data <- 
  bind_rows(obs, dose) %>% 
  arrange(ID, time) %>%
  mutate(Subject = NULL)


head(data)
```


- Take `dv` as a vector of observed concentrations

```{r}
dv <- obs$conc
```


# Create an objective function function

- For starters, just do OLS estimation
- Note that we _need_ to name the parameters `p`
    - Parameter updates require names in `mrgsolve`
    - Generally, don't expect `p` to retain any names that you might
    pass in through the initial estimates
- We also pass in the `data` and the dependent variable (`dv`)

```{r}
obj <- function(p, theta, data, dv, pred = FALSE, obsonly = TRUE) {
  
  names(p) <- names(theta)
  
  p <- lapply(p,exp)
  
  mod <- param(mod, p)
  
  out <- mrgsim(mod, data = data, obsonly=obsonly)
  
  if(pred) return(as.data.frame(out))
  
  return(sum(((out$CP - dv))^2))
}
```

# Load a PK model 

- We'll try out one-compartment first
```{r}
mod <- mread_cache("pk1", modlib())

param(mod)
```

# Fit with one-compartment model

- First generate some initial estimates
- These _need_ to be named in a way that is consistent with
the model we are using
- I usually run a test with the objective function function to 
make sure the logic works out

```{r}
theta <- log(c(CL = 1, V = 100))

obj(theta,theta,data,dv)
```

- Nelder-Mead optimization
```{r}
fit <- optim(par = theta, fn=obj, theta = theta, data=data, dv=dv)
```

- And generate some predictions based on the final estimates
```{r}
pred <- obj(fit$par, theta, data, dv, pred = TRUE, obsonly = FALSE)

data$pred <- pred$CP

head(data)
```

# Make a plot of the output

- What do you think?  Good fit?

```{r}
ggplot(data = data) + 
  geom_point(aes(time,conc)) + 
  scale_y_continuous(trans="log") + 
  geom_line(aes(time,pred),col="firebrick")
```

# Your turn

- Load a 2 compartment model


```{r}
mod <- mread_cache("pk2", modlib())
```


- Take a look at the model and generate a call to `minqa::newuoa` using 
the OLS objective function above to fit the data

- You will also need try out a new set of initial estimates for all of
the volumes and clearances for 2-compartment, IV bolus model

- What do you think of the fit using the the OLS objective function?
    - Can you make a simple modification to the OLS objective function
    that might make the fit look a little better?

- Suppose we're worried about the `newuoa` optimizer and want to 
try a global search algorithm
    - Can you construct a call to `RcppDE::DEoptim` that will also 
    fit the data?
    - Remember that `DEoptim` doesn't use initial estimates the 
    same way `stats::optim` or `minqa::newuoa` does; you have to 
    specify one vector of lower boundaries and one vector of 
    upper boundaries, with a lower and upper bound for each parameter



# Answer

- Set the initial estimates for two compartment model

```{r}
param(mod)

theta <- log(c(CL = 2, V2 = 4, Q = 5, V3 = 10))
```

- Try weighted least squares 

```{r}
obj <- function(p, theta, data, dv, pred = FALSE, obsonly = TRUE) {

  names(p) <- names(theta)
  
  p <- lapply(p,exp)
  
  mod <- param(mod, p)
  
  out <- mrgsim(mod, data = data, obsonly=obsonly)
  
  if(pred) return(as.data.frame(out))
  
  return(sum(((out$CP - dv)/dv)^2))
}
```

```{r}
fit <- minqa::newuoa(par = theta, fn=obj, theta = theta, data=data, dv=dv)
```

Final estimates and final value of objective function
```{r}
exp(fit$par)

obj(fit$par,theta,data,dv)
```

- Generate predictions for the final and initial estimates

```{r}
pred <-  obj(fit$par, theta, data, dv, pred = TRUE, obsonly = FALSE)
predi <- obj(theta,  theta, data, dv, pred = TRUE, obsonly = FALSE)


data$pred <- pred$CP
data$predi <- predi$CP
head(data)
```

- Plot the predictions
```{r}
pred <- distinct(data, time, .keep_all = TRUE)

ggplot(data = data) + 
  geom_point(aes(time,conc)) + 
  scale_y_continuous(trans="log") + 
  geom_line(data=pred,aes(time,pred),col="firebrick", lwd=1) +
  geom_line(data=pred,aes(time,predi),col="darkgreen", lwd=1) 
```

## Fit the data with `RcppDE::DEoptim`

```{r, eval = FALSE}
fit <- RcppDE::DEoptim(obj, lower = rep(-4,4), upper = rep(4,4), 
                       theta = theta, data = data, dv = dv)

```

## Check the estimates and the final value of the objective function

```{r, eval = FALSE}
exp(fit$optim$bestmem)

obj(fit$optim$bestmem, theta,data,dv)
```

