---
title: ""
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: docs/src/styles/styles.css
---

# Quantitative Analyses of Hepatic OATP-Mediated Interactions Between Statins and Inhibitors Using PBPK Modeling With a Parameter Optimization Method


```{r echo=FALSE, message=FALSE}
source("libPaths.R")
library(mrgsolve)
library(dplyr)
library(ggplot2)
options(mrgsolve_mread_quiet=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, comment='.')
```

<hr><div class = spacer></div>

# Cyclosporine simulation
```{r}
mod <- mread("yoshikado", "model")
ec <- ev(amt=2000, cmt=2)
out <- mod %>% ev(ec) %>% mrgsim(delta=0.01,end=12)
sims <- as.tbl(out)
```

```{r}
ggplot(sims, aes(x=time,y=CSA)) + 
  geom_line() + 
  scale_y_continuous(trans="log", limit=c(50,5000),breaks=10^seq(-4,4)) + 
  scale_x_continuous(breaks=seq(0,12,2)) 
```

<hr><div class = spacer></div>


# Pitavastatin (EHC model) simulation

- Single 30 mg dose of pitavastatin

```{r}
ep <- ev(amt=30)
out <- mod %>% ev(ep) %>% mrgsim(delta=0.1,add=seq(0,0.5,0.01), end=21)
simsp <- as.tbl(out)
```


```{r}
ggplot(simsp, aes(x=time,y=CP)) + 
scale_y_continuous(trans="log10", limit=c(0.1,100),breaks=c(0.1,1,10,100)) + 
geom_line() + scale_x_continuous(breaks=seq(0,14,2),limit=c(0,14)) 
```

<hr><div class = spacer></div>

# Now, simulate the DDI
```{r}
out <- mod %>% ev(ep + ec) %>% mrgsim(delta=0.01, end=14)
```


## Blue: pitavastatin, no ddi; red: pitavastatin with CsA ddi
```{r}
simspc <- as.tbl(out)
ggplot(simspc, aes(x=time,y=CP)) + 
  scale_y_continuous(trans="log10", limit=c(0.1,100), breaks=c(0.1,1,10,100)) + 
  geom_line(col="firebrick") + scale_x_continuous(breaks=seq(0,14,2),limit=c(0,14)) + 
  geom_line(data=simsp, aes(x=time,y=CP),col="darkslateblue")
```

<hr><div class = spacer></div>

# Sensitivity analysis

```{r}
library(MASS)
```


Pick some parameters for sensitivity analysis

```{r}
param(mod)
l <- as.numeric(param(mod))
l
```

First, look at partition coefficients
Maybe try the flows later

```{r}
p1 <- grepl("^Q", names(l))
p2 <- grepl("Kp_", names(l))
l <- l[p2]
l
```

Run the sensitivity analysis with idata_set
```{r}
set.seed(81911)

idata <- mvrnorm(500,log(l),diag(rep(1, length(l))))

idata %<>% exp %>% as.data.frame %>%  mutate(ID = 1:n())

head(idata)
```

```{r}
sims <- 
  mod %>% ev(ep+ec) %>%
  obsonly %>% Req(CP) %>%
  mrgsim(delta=0.25,add=0.025,end=14,idata=idata) %>%
  filter(time > 0)
```


```{r}
dim(sims)
head(sims)
```


```{r}
ggplot(sims, aes(x=time,y=CP, group=ID)) + 
  scale_y_continuous(trans="log10", limit=c(0.1,300), breaks=c(0.3,3,30,300)) + 
  geom_line(col="darkslateblue") + 
  scale_x_continuous(breaks=seq(0,14,2),limit=c(0,14)) +
  geom_line(data=simspc, col="orange", lwd=1) + 
  geom_line(data=simsp, col="firebrick", lwd=1)  
```

<hr><div class = spacer></div>

# Sensitivity analysis on inhibitor Ki for CSA

```{r}

N <- 300

set.seed(8811112)

idata <- data_frame(ID=1:N, ikiu = rnorm(N,log(mod$ikiu),sqrt(0.5)))

idata %<>% exp %>% as_data_frame %>% mutate(ID = 1:n())

idata
```


```{r}
sims <- 
  mod %>% ev(ep+ec) %>%
  obsonly %>% Req(CP) %>% carry_out(ikiu) %>%
  mrgsim(delta=0.1,add=0.025,end=14,idata=idata) %>%
  filter(time > 0)
dim(sims)
head(sims)
```

* `red` no DDI
* `orange` DDI, `ikiu` at default value
* `blues` sensitivity analysis on `ikiu`

```{r}
ggplot(sims, aes(x=time,y=CP, group=ID)) + 
  scale_y_continuous(trans="log10", limit=c(0.1,300), breaks=c(0.3,3,30,300)) + 
  geom_line(aes(col=ikiu)) + 
  scale_x_continuous(breaks=seq(0,14,2),limit=c(0,14)) +
  geom_line(data=simspc, col="orange", lwd=1)  + 
  geom_line(data=simsp, col="firebrick", lwd=1)  
```


<hr><div class = spacer></div>


# mrgsolve tool kit


```{r}
library(mrgsolvetk)
```


https://github.com/mrgsolve/mrgsolvetk


```{r}
out <- 
  mod %>% 
  ev(ep+ec) %>% update(delta = 0.1, end = 14) %>% Req(CP) %>%
  sens_norm(n = 25, pars = names(l), cv = 100)
out
```


```{r}
ggplot(as.data.frame(out), aes(time,CP, col=factor(ID))) + 
  geom_line() + scale_y_continuous(trans = "log10")
```







