---
title: "Clinical trial simulation with `mrgsim`"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: docs/src/styles.css
---

<hr><div class = spacersm></div>

```{r, echo = FALSE}
knitr::opts_chunk$set(comment='.', warning=FALSE, message=FALSE)
```

# Packages
```{r}
library(mrgsolve)
library(mrgsim)
library(ggplot2)
library(dplyr)
```

<hr><div class = spacer></div>

# PK / PD model
```{r}
mod <- mrgsolve:::house() %>% omat(dmat(0.1, 0.1, 0.1, 0.1))


mod %>% 
  ev(amt = 1000, rate = 50, cmt = 2, time = 10) %>% 
  mrgsim %>% plot(CP + RESP ~.)
```

<hr><div class = spacer></div>

# The `mrgsim` package

Allows you to lay out treatment sequences, arms, populations, designs
for a clinical trial in `yaml` format

```{r}
writeLines(readLines("mrgsim.yml"), sep="\n")
```



```{r}
x <- load_run("mrgsim.yml")

arms(x)
```

```{r}
out <- sim_run(mod, x)
```

```{r}
filter(out, time ==0) %>% count(arm) %>% left_join(.,arms(x))
```


<hr><div class = spacer></div>

Take one `ID` from each arm and plot `CP` versus time to see what happened

## Trough concentration over time (arm1)
```{r}
out %>% filter(ID ==1) %>% ggplot(., aes(time,CP)) + geom_line()
```


## Trough concentration over time (arm2)
```{r}
out %>% filter(ID ==260) %>% ggplot(., aes(time,CP)) + geom_line()
```
## Trough concentration over time (arm3)
```{r}
out %>% filter(ID ==600) %>% ggplot(., aes(time,CP)) + geom_line()
```

<hr><div class = spacer></div>

## Summary of trough response by treatment arm
```{r}
filter(out, time==4032) %>% 
  group_by(arm,armn) %>%
  summarise(resp = mean(RESP), sd = sd(RESP))
```


<hr><div class = spacer></div>

