---
title: "MAPK"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: docs/src/styles/styles.css
    toc: true
    number_sections: true
---

```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE, fig.width=8)
```

__Clinical responses to ERK inhibition in BRAF{V600E}-mutant colorectal cancer predicted using a computational model__

- Daniel C. Kirouac, Gabriele Schaefer, Jocelyn Chan, Mark Merchant, Christine Orr, Shih-Min A. Huang, John Moffat, Lichuan Liu, Kapil Gadkar and Saroja Ramanujan

- npj Systems Biology and Applications  (2017) 3:14 

- https://www.ncbi.nlm.nih.gov/pubmed/28649441

```{r}
a <- config::get()
library(mrgsolve)
library(readr)
library(purrrlyr)
library(purrr)
library(dplyr)
library(dmutate)
library(ggplot2)
library(magrittr)
source("src/functions.R")
set.seed(889110)
```

```{r}
mod <- mread("mapk", "model")
see(mod)
```



# Read in the VPOP data set provided by the authors

- This is the data file that came with the publication
```{r}
data.file <- file.path("data", "s10vpop.csv")
```

- Just sample 250 rows, with the weights provided in the `PW` column
```{r}
set.seed(1001100)
vp <- read_csv(data.file) %>% sample_n(250, weight=PW, replace=TRUE)
```

Since we sampled with replacement, let's add a new VPOP column that is unique per row
```{r}
vp %<>% mutate(VPOP2 = 1:n())
```


# Read in PK parameters 

- There are a lot of compounds to deal with
- I've arranged the PK parameters in a list that will help keep us compatible with `dmutate`
```{r}
source("src/objects.R")

names(pki)
pki$iiv_cobi

pki$cl1
```

# Use `dmutate` package to simulate PK parameters

- Simulate from a multivariate normal distirbution, one per unique `VPOP2`
- These parameters are for `RAFi`
```{r}
cov2 <- cl2 + ka2 + V2 ~ rlmvnorm(log(c(cl2,ka2,V2)),iiv_vemu) | VPOP2

```

- Similar deal for `MEKi` and `ERKi`
```{r}
cov3 <- cl3 + ka3 + q2 + V3 + V3b ~ 
  rlmvnorm(log(c(cl3,ka3,q2,V3,V3b)),iiv_cobi) | VPOP2

cov4 <- cl4 + ka4 + V4 ~ 
  rlmvnorm(log(c(cl4,ka4,V4)),iiv_gdc) | VPOP2
```

- We collect these objects in to a `covset` ... 
```{r}
cov_set <- covset(cov2,cov3,cov4)

cov_set
```

- And we can mutate the `vp` data set to add the simulated PK values
```{r}
vp %<>% mutate_random(cov_set, envir = pki)
#vp <- mutate_random(vp, cov_set, envir = pki)

names(vp)
```

- Derive the rate constants; we must do this matching names in the model
```{r}
vp %<>% mutate(ke2 = cl2/V2, ke3 = cl3/V3, ke4 = cl4/V4)
```


Now, split up the Vpop in to a list that we can use to simulate
```{r}
l <- split(vp, vp$VPOP2)
length(l)
```

# Create all mono-tx regimens that we will need to address 

- Nothing
```{r}
data0 <- expand.ev(amt=0, cmt=8)
```

- BRAF-i CMT 8  - vemurafanib VEMU
- VEMU_dose = 960;    %mg, BID
```{r}
cmtn(mod, "RAFi_gut")
dataV <- expand.ev(amt=960, cmt=8, ii=0.5, addl=120)
```

- ERKi CMT 12 - GDC-0994 
- GD ERKI_dose = 400;    %mg, daily 3/1
```{r}
cmtn(mod, "ERKi_gut")

dataG <- 
  expand.ev(amt=400, cmt=12, time=c(0,28), ii=1, addl=20) %>% 
  mutate(ID=1)

dataG
```

- Another way to make this regimen
```{r}
e <- ev(amt = 400, cmt = 12, ii = 1, addl = 20, ID = 1)

datag <- ev_repeat(e, n = 2, wait = 7)
datag
```

- Give it a test
- This plot shows the trough concentration versus time 
```{r}
out <- mrgsim(mod, data=datag, end=56)
plot(out, ERKi~time)
```


- MEKi CMT 10 - cobimetinib COBI
- COBI_dose = 60;     %mg, daily
```{r}
cmtn(mod,"MEKi_gut")
dataCO <- mutate(dataG, amt=60, cmt=10)
```

- RTK1i / EGFRi CMT 7 - cetuximab CETUX
- CETUX_dose = 450;   %mg, weekly
```{r}
cmtn(mod, "RTK1i_blood")

dataCE <- data_frame(time=0, cmt=7, ii=7, addl=7, evid=1, ID=1, amt=450)
```

# A function that combines two or more regimens

- We already have all of the individual regimens
- The function combines any number of regimens that we
pass in to create the combinations

```{r} 
comb <- function(...) {
  bind_rows(list(...)) %>% dplyr::arrange(time)
}
```

- Try combining the cetuximab and vemurafanib regimens
```{r}
comb(dataCE, dataV)
```


```{r}
mod %>% 
  mrgsim(data = comb(dataCE,dataV), end = 14, delta = 0.1) %>%
  plot(RTK1i_blood + RAFi_blood + TUMOR ~ time)
```


# The simulation function

- This function is designed to let us simulate from one element in the 
`vpop` list
- We pass in one row from the `vpop` data set, update parameters 
and initial conditions, and simulate with the data set that was also passed in


```{r}
sim <- function(x, Data) {
  mod %>%
    param(x) %>%
    init(x) %>% 
    Req(TUMOR) %>%
    mrgsim(data=Data, end=56, delta=0.25) %>%
    filter(time==56)
}
```

- I frequently also create a function that will simulate the 
whole list of stuff
- So `siml` calls `sim` for each position in the list
- This is also a good place to do some post-processing or modification 
of the output with some information that we might need later

```{r}
siml <- function(data) {
  parallel::mclapply(l, mc.cores=8, function(x) {
    sim(x,Data=data) %>% 
      mutate(VPOP2=x$VPOP2,VPOP=x$VPOP)
  }) %>% bind_rows
}
```


# Simulate all of the regimens

- I'm going with brute-force approach here
- Maybe more elegant approach is possible
- But often brute-force is more flexible, easier to debug,  or 
easier to re-do this or that scenario


## Monotherapy
```{r}
sim1 <- siml(data0 ) %>% mutate(label=1)
sim2 <- siml(dataCE) %>% mutate(label=2)
sim3 <- siml(dataV ) %>% mutate(label=3)
sim4 <- siml(dataCO) %>% mutate(label=4)
sim5 <- siml(dataG ) %>% mutate(label=5)
```

## 2-drug combinations
```{r}
sim23 <- comb(dataCE, dataV)  %>% siml %>% mutate(label=23)
sim24 <- comb(dataCE, dataCO) %>% siml %>% mutate(label=24)
sim25 <- comb(dataCE, dataG)  %>% siml %>% mutate(label=25)
sim34 <- comb(dataV,  dataCO) %>% siml %>% mutate(label=34)
sim35 <- comb(dataV,  dataG)  %>% siml %>% mutate(label=35)
sim45 <- comb(dataCO, dataG)  %>% siml %>% mutate(label=45)
```

## 3-drug combinations
```{r}
sim234 <- comb(dataCE, dataV,  dataCO) %>% siml %>% mutate(label=234)
sim235 <- comb(dataCE, dataV,  dataG)  %>% siml %>% mutate(label=235)
sim245 <- comb(dataCE, dataCO, dataG)  %>% siml %>% mutate(label=245)
sim345 <- comb(dataV,  dataCO, dataG)  %>% siml %>% mutate(label=345)
```

## All 4 drugs
```{r}
sim2345 <- comb(dataCE, dataV, dataCO, dataG) %>% siml %>% mutate(label=2345)
```

# Process and plot
```{r}
lab <- c("No TREAT", "CETUX", "VEMU", "COBI", "GDC",
         "CETUX+VEMU", "CETUX+COBI", "CETUX+GDC", "VEMU+COBI","VEMU+GDC",
         "COBI+GDC",
         "CETUX+VEMU+COBI", "CETUX+VEMU+GDC", "CETUX+COBI+GDC", "VEMU+COBI+GDC",
         "CETUX+VEMU+COBI+GDC")
```

```{r}
sims <- bind_rows(sim1,sim2,sim3,sim4,sim5,sim23,sim24,sim25,sim34,
                  sim35,sim45,sim234,sim235,sim245,sim345,sim2345)
```

```{r}
ulab <- unique(sims$label)
sims %<>% mutate(labelf = factor(label,levels=ulab,labels=as.character(ulab)))
sims %<>% mutate(labelff = factor(label,levels=ulab,labels=lab))
```


```{r}
p1 <- 
  ggplot(data=sims) + 
  geom_point(aes(x=labelff, y=TUMOR),position=position_jitter(width=0.15),col="grey") +
  scale_y_continuous(limits=c(0,2.5),name="Tumor size",breaks=c(0,0.5,1,1.5,2,2.5,3)) +
  scale_x_discrete(name="") + 
  geom_hline(yintercept=0.7,col="firebrick", lty=1,lwd=1)  +
  geom_boxplot(aes(x=labelff, y=TUMOR),fill="darkslateblue",col="darkslateblue",alpha=0.2) +
  theme_plain() + rotx(30)
```

```{r}
p1
```


```{r}
pdf(file="all_treat_.pdf", width=8,height=4)
p1
dev.off()
```
