---
title: "Plugin"
output: 
  html_document:
    css: docs/src/styles/styles.css
    toc: true
    number_sections: true
---

```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE)
```

# Objective

This vignette will introduce plugins that can be used with mrgsolve.

# Setup
```{r}
library(mrgsolve)
library(dplyr)
```


# PLUGIN

* Recruit extra functionality in your `mrgsolve` model
    - Random variates 
    - Complex data structures
    - Call `R` functions
    - Call `C++` functions from other libraries (like `boost`)
* Via `Rcpp` or `C++` coding
* By definition require extra libraries to be compiled into the model
    - The plugin mechanism handles all of the details
* We make this a plugin because most models still don't require them
    - We're introducing this topic because more and more we need this functionality
  
  
# The `Rcpp` plugin

* The most basic plugin
  * Without this plugin, the __model__ is implemented in pure `C++`
* Gives you access to `Rcpp` vectors and matrices and `R` functions
* Gives you access to `Rcpp::R` namespace


## Random number generation with `Rcpp`


```{r}
code <- '
$PLUGIN Rcpp

$MAIN
capture y = R::rnorm(10,2);
'
```

* We use `$PLUGIN` code block and request the `Rcpp` plugin
* `Rcpp` provides a `C++` namespace called `R`
* The `R` namespace includes all the expected `dpqr` functions
  - http://dirk.eddelbuettel.com/code/rcpp/html/namespaceR.html


This takes a little longer to compile
```{r}
mod <- mcode_cache("rnorm", code)
```

```{r}
mrgsim(mod) 
```

Now, we have a random variate called `y` in the model 
that is normally distributed with some mean and 
some variance.  

This was just a demo to get the mechanics down.  Let's 
imagine we want to code some covariates for each subject 
in the simulation.


```{r}
code <- '
$PLUGIN Rcpp

$PARAM TVCL = 20, CL_FEMALE = 0.85, pfemale = 0.6

$MAIN
if(NEWIND <= 1) {
  capture WT = R::rnorm(70,12);
  capture SEX = R::rbinom(1,pfemale);
}

capture CL = TVCL * pow(WT/70, 0.75);
if(SEX==1) CL = CL * CL_FEMALE;
'
```

```{r}
mod <- mcode_cache("rnorm_wt", code)
```


```{r}
mrgsim(mod)
```


We probably should have set a seed ...

```{r}
mrgsim(mod, end=-1, nid=5) 
```


### <alrt>TRY it</alrt>

* Name: <dsb>Mixture model</dsb>
* Create a model specification file that simulates a mixture model
* Start with a clearance difference in two populations
* Next, try to code a clearance difference in three populations


## Call an R function

* It is entirely possible to call `R` functions from your
`mrgsolve` model
* Sometimes it is helpful to accomplish your objectives
* But note well that there is a performance ding every time
you go back into `R` to call a function; the ding doesn't 
mean the simulation time will increase x10, but be mindful 
of the ding and be wise when going to this option
* I would be much more likely to call the `R` function
once per individual or once at the start of the problem;
you should almost never call the `R` function inside `$ODE`

Here we go ... Implement a simple reporter functionality 

* In `$GLOBAL`, we declare an `Rcpp::Function` that is
conveniently called `print`


```{r, eval = FALSE}
code <- '

$PLUGIN Rcpp

$GLOBAL
Rcpp::Function print("print");


$MAIN
double y = R::rnorm(100,10);

print(y);
'
```

```{r, eval = FALSE}
mod <- mcode_cache("print", code)

out <- mrgsim(mod, end=10)
```

# `mrgx` plugin

This plugin provides some functionality that 
you might find helpful. 

## Simulation of random variates inside a lower and upper limit:

```{r, eval = FALSE}

Sys.sleep(1)

code <- '

$PLUGIN mrgx

$MAIN 
if(NEWIND <= 1) {
  double WT = mrgx::rnorm(70, 70, 60, 80);
}

$CAPTURE WT
'
```


```{r}
mod <- mcode_cache("mrgx", code)

mrgsim(mod, end=-1, nid = 100)
```



## <alrt>TRY it</alrt>

* Name: <dsb>clin trial</dsb>
* Create a self-contained clinical trial simulator








