---
title: "Simulate from Emperical Bayes Estimates"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: docs/src/styles/styles.css
    toc: true
---

# Packages

```{r,message = FALSE}
source("libPaths.R")
library(readr)
library(dplyr)
library(mrgsolve)
library(magrittr)

a <- config::get()
source(a$functions)
knitr_config()
```


# Objective
for each individual in the study, get the 
model-predicted Cmin, Cmax, and AUC at steady state
for the randomized dose administration

- We use empirical Bayes estimates that come out 
of the NONMEM run
- There are multiple doses and infusion durations represented
in the data; we need to get the right doses to the right 
individuals
- Simulate out 4 doses should be enough
- Calculate AUC by trapezoidal rule; use the 
parametric AUC calculation as a QC check




# Load input data

## Load the meropenem partab file

```{r}
tab <- read_table(a$mero_partab, skip=1)
```


We create a data frame with one ID per row including each 
individual's PK parameters
```{r}
idata <- distinct(tab,ID,CL,V1,Q,V2)
```

## Load the meropenem data set

```{r}
data <- read_csv(a$mero_datafile, na='.')
names(data)[1] <- "ID"
```


And do some grooming
- Take only dosing records
- meropenem is dosing every 8 hours
- simulate a total of 12 doses
- for IV administration, pick compartment 2

```{r}
dose <- filter(data, EVID==1)

dose %<>% mutate(II=8, ADDL=11, CMT=2)
```


```{r}
count(dose,AMT,RATE)
```

## Join the data set with the individual parameters

```{r}
data <- left_join(dose,idata)

head(data)
```



This is important: we derive a `DOSE` column
so that we can work with the simulated data according 
to dose group later on

```{r}
data %<>% mutate(DOSE = AMT, DUR = AMT/RATE)
```

# Load a 2-cmt model

Just take a model from the library

```{r}
mod <- mread("pk2", modlib())

see(mod)
```


Check the parameters for this model

```{r}
param(mod)
```


We should inventory the parameters in the model against 
what is in the data set

```{r}
inventory(mod,data)
```



Easy enough to fix
```{r}
data %<>% rename(V3 = V2, V2 = V1) %>% mutate(V1 = NULL)

inventory(mod,data)
```


The model we are using includes a extravascular dosing 
compartment with first order absorption controlled by 
parameter `KA`.  Our drug is administered `IV` so 
it's ok that we're missing `KA` in the data set.

Not required ... but I update the model with the 
parameters from the first individual to have a set 
of "valid" parameters in the model.  Sometimes this
helps when trying to get a sense check about whether 
the simulation is set up correctly or not.

```{r}
mod %<>% update(param=data[1,])
```



Simulate from a time grid out to the 3rd day
```{r}
des <- tgrid(0, 24, 0.5) + 3*24
```



In the `mrgsim` call, let's ask for `DOSE`, `DUR` and `CL` in 
the simulated output.  
```{r}
count(data,DOSE,AMT,DUR)
```



What would happen if we asked for `AMT` instead of `DOSE` here?

```{r}
out <- mrgsim(mod, carry.out="DOSE,DUR,CL",
              data=data, tgrid=des, obsonly=TRUE)
```

```{r}
head(out)

plot(out,CP~.)
```


We get a mix here. You can see the variability due to duration of 
the infusion.  It's not so clear who had the `500`, `1000`, and `5000` mg doses.

So, let's summarize.

```{r}
sims <- as_data_frame(out)
```

Calculate `AUC` by trapezoidal rule
```{r}
auc <- 
  sims %>% 
  AUC(., dv = "CP") %>% 
  dplyr::rename(tAUC = AUC) %>%
  as_data_frame
```


```{r}
auc
```

Recall that we carried `CL` into the simulated data. Let's calculate `AUC` through
`CL` and `DOSE` to see how it matches up

```{r}
auc2 <- 
  sims %>%
  distinct(ID,DOSE,CL) %>%
  mutate(pAUC = 3*DOSE/CL) %>% 
  ungroup

auc2
```


Calculate `Cmin` and `Cmax`
```{r}
minmax <- sims %>% 
  group_by(ID,DOSE,DUR) %>% 
  summarise(Cmin = min(CP), Cmax = max(CP)) %>% 
  ungroup

minmax
```




Join together and re-shape

```{r}
sum <- left_join(auc, minmax) %>% left_join(.,auc2)

summ <- tidyr::gather(sum, variable, value, c(tAUC,pAUC,Cmin,Cmax))

```


# Summary of exposure metrics
```{r}
summ %>% 
  ungroup %>%
  group_by(DOSE,DUR,variable) %>%
  summarise(Med = median(value), 
            Mean = mean(value),
            Min = min(value),
            Max = max(value),
            N=n())  %>% as_data_frame
```

