---
title: "Visual predictive check - EPO model"
author: "Metrum Research Group, LLC"
date: ""
output: 
  github_document: 
    toc: true
---

```{r, echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, warning=FALSE)
```


```{r}
library(dplyr)
library(mrgsolve)
library(readr)
library(magrittr)
library(ggplot2)
library(parallel)
```


# Some helper functions
```{r}
qt <- function(x,y) unname(quantile(x,prob=y/100))
lo <- function(x) qt(x,5)
hi <- function(x) qt(x,95)
med <- function(x) qt(x,50)
loci <- function(x) qt(x,2.5)
hici <- function(x) qt(x,97.5)
col1 <- "steelblue"
col2 <- "firebrick"
```

Not really ground-breaking stuff.  But the code might get a little
messy later on and I'd like to have something more
readable if I can.

# Model estimation data set
```{r}
a <- config::get()
data <- read_csv(a$epo_datafile) %>% as_data_frame
```


```{r}
head(data)
```


Notice that the data set has upper case names.  So, when we're adding
columns let's make them upper case too.


We need to add an `EVID` column

```{r}
data %<>% mutate(EVID = as.integer(AMT > 0))
```


Add a flag for doses that were given subcutaneously
```{r}
data %<>% 
  group_by(ID) %>% 
  mutate(SC = as.integer(any(CMT==1))) %>%
  ungroup

data
```



# Compartments

1. EPO SC
2. EPO CENTRAL
7. RETICULOCYTES
8. RED BLOOD CELLS
9. HEMOGLOBIN

```{r}
count(data,CMT,EVID)
```

Just some grooming

```{r}
data %<>% group_by(ID,TIME,CMT,EVID) %>% slice(1) %>% ungroup

count(data,CMT,EVID)

data0 <- data
```


Take reciculocyte measurements after `SC` epo administration
```{r}
data <- filter(data, 
               CMT==7 | (EVID==1 & CMT %in% c(1,2)), 
               TIME <= 672 & SC==1)
```


I'm setting a nominal time here for plotting 
```{r}
data %<>% mutate(NTIME = round(TIME))
```



Isolate the observation records from the dosing records

```{r}
obs <- filter(data, EVID==0 & DV !=0)
dose <- filter(data, EVID==1)
```


Preview the observations
```{r}
ggplot(data=obs) + geom_point(aes(TIME,DV)) 
```



# Load the epo model
```{r}
mod <- mread("epo", a$epo_project)
mod
```

# Simulate

Just going a handlful of iterations for now

```{r}
niter <- 50
```

```{r}
simvpc <- function(i) {
  mod %>% 
    data_set(data) %>%
    obsonly %>%
    carry_out(NTIME) %>%
    mrgsim() %>% 
    mutate(irep = i)
}
```


Summarize the observations by nominal time
```{r}
sumo <- 
  obs %>% ungroup %>%
  group_by(NTIME) %>% 
  summarise(med = med(DV), lo=lo(DV), hi=hi(DV), N = n()) %>%
  filter(N > 10) 
```


Simulate using `parallel::mclapply`

```{r}
library(parallel)
out <- mclapply(1:niter, mc.cores=4, simvpc) %>% bind_rows
```



Now,summarize the simulated data by nominal time
```{r}
sum1 <- 
  out %>% 
  group_by(irep,NTIME) %>%
  summarise(med = med(RETdv), lo = lo(RETdv), hi=hi(RETdv), N=n())
```



And summarize the summaries
```{r}
sum2 <- 
  sum1 %>% 
  filter(N > 10) %>%
  group_by(NTIME) %>% 
  summarise(medlo = loci(med), medmed = med(med), medhi = hici(med),
            lolo =  loci(lo),  lomed  = med(lo),  lohi  = hici(lo),
            hilo =  loci(hi),  himed  = med(hi),  hihi  = hici(hi))
```




# VPC plot

I usually end up using brute force for this; but there 
may be more-elegant ways to melt and plot.  In our experience,
it's easier to QC this way

```{r, fig.width=8}
ggplot(data=sum2) + 
  geom_ribbon(aes(NTIME,ymin=medlo, ymax = medhi),alpha=0.3,fill=col1) + 
  geom_ribbon(aes(NTIME,ymin=lolo, ymax=lohi),alpha=0.3,fill=col1) + 
  geom_ribbon(aes(NTIME,ymin=hilo, ymax=hihi),alpha=0.3,fill=col1) + 
  geom_point(data=obs, aes(TIME,DV),col=col2) +
  geom_line(aes(NTIME,y=medmed), lwd=1,col=col1) +
  geom_line(aes(NTIME,y=lomed), lwd=1,col=col1) +
  geom_line(aes(NTIME,y=himed), lwd=1,col=col1) +
  geom_line(data=sumo, aes(NTIME,med),lty=2,lwd=1) + 
  geom_line(data=sumo, aes(NTIME,lo),lty=2,lwd=1) +
  geom_line(data=sumo, aes(NTIME,hi),lty=2,lwd=1) 
```


  
 

