---
title: "Parameter estimation in PBPK model"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document: 
    toc: true
    number_sections: true
    css: styles/styles.css
---

```{r, echo = FALSE}
knitr::opts_chunk$set(comment = '.', warning = FALSE, message = FALSE)
```

# Packages
```{r packages}
a <- config::get()

library(readr)
library(dplyr)
library(mrgsolve)
library(minqa)
library(magrittr)
library(tidyr)
library(ggplot2)
source(a$functions)
```

# Definitions
```{r definitions }
typef <- function(x) {
  factor(x, c(1,2), c("Pitavastatin alone", "Pitavastatin + CsA"))
}
```

# Input data set
```{r read data}
data <- read_csv(a$fit_03_data) %>% 
  mutate(profile = NULL,type=ID,typef=typef(ID))

head(data)
```


```{r data preview}
ggplot(data=data,aes(time,DV)) + 
  geom_point(col="firebrick") + 
  facet_wrap(~typef) + 
  scale_y_continuous(trans="log", limits=c(0.1,300), breaks=logbr())
```


```{r}
dose <- filter(data,evid==1) %>% mutate(typef=NULL)

yobs <- filter(data,evid==0) %>% dplyr::select(DV) %>% unlist %>% unname
```

```{r weighted sum of squares}
wss <- function(dv,pred,par=NULL) sum(((dv-pred)/dv)^2)
```

```{r read in model}
mod <- mread("yoshikado",a$ddi_project) %>% 
  update(end=14,delta=0.1) %>% Req(CP) %>% obsonly
```

```{r}
data %<>% dplyr::select(-typef)

mod %>% 
  mrgsim(data=dose,obsaug=TRUE) %>% 
  plot(CP~.,scales=list(y=list(log=TRUE)))
```

# Prediction function

- Arguments
    - `p` parameters on *this* iteration
    - `.data` the data set
    - `yobs` the dependent variable
    - `pred` logical
- First, untransform the parameters
- Then update the paramter list
- Simulate
- Return the weighted sum of squares

```{r pred function}
pred <- function(p, .data, yobs=NULL, pred=FALSE) {
  
  .par <- lapply(p,exp) %>% setNames(names(theta))
  
  .mod <- param(mod,.par)
  
  if(pred) {
    out <- mrgsim(.mod, data=.data, carry.out="type")
    return(as_data_frame(out))
  }
  
  out <- mrgsim(.mod, data=.data, obsonly=TRUE, Req="CP")
  
  return(wss(yobs,out$CP))
  
  #return(-1*sum(dnorm(log(yobs),log(out$CP),.par$sigma,log=TRUE)))
  
}
```

# Fit with newuoa

- 5 parameters on log scale

```{r}
theta <- log(c(fbCLintall = 1.2, ikiu = 1.2, 
               fbile = 0.9, ka = 0.1, ktr = 0.1))

control <- list(iprint=25)

fit <- newuoa(theta, pred,.data=data, yobs=yobs, control=control)
```

```{r}
fit$par <- setNames(fit$par,names(theta))
```

## Simulate data for prediction lines

- First the predicted with final estimates
- Then the predicted with initial estimates
- Also get a set of observed data ready for the plot
```{r}
df_pred <- pred(fit$par, dose, pred=TRUE) %>% mutate(type=typef(type))
df_init <- pred(theta,   dose, pred=TRUE) %>% mutate(type=typef(type))
df_obs <- mutate(data, type=typef(type))
```

## Plot
```{r}
ggplot(data=df_pred) + 
  geom_line(data=df_init,aes(time,CP,lty="A"), col="black", lwd=0.7) +
  geom_line(aes(time,CP,lty="B"),col="darkslateblue",lwd=0.7) + 
  geom_point(data=df_obs,aes(time,DV),col="firebrick",size=2) + 
  facet_wrap(~type) + 
  scale_y_continuous(trans="log",breaks=10^seq(-4,4), 
                     limits=c(0.1,100),
                     "Pitavastatin concentration (ng/mL)") +
  scale_x_continuous(name="Time (hours)", breaks=seq(0,14,2)) +
  scale_linetype_manual(values= c(2,1),labels=c("Initial estimates", "Final estimates"), name="") +
  theme(legend.position="top")
```


## Final OFV value

```{r}
pred(fit$par,.data=data,yobs=yobs)
```

## Final estimates
```{r}
exp(fit$par)
```

# Use `optimhelp` to manage parameters
```{r}
library(mrgsolvetk)
library(optimhelp)
```

```{r}
par <- parset(log_par("fbCLintall",1.2),
              log_par("ikiu", 1.2), 
              logit_par("fbile", 0.8), 
              log_par("ka", 0.1),
              log_par("ktr", 0.1))
```

```{r}
fitt <- fit_optim(mod,
                  data, pred="CP", ofv=wss,
                  par=par,method="CG",
                  control=list(trace=10))
```

```{r}
coef(fitt$pars) 
```

```{r}
fitt$value
```



# DEoptim
- "Performs evolutionary global optimization via the Differential Evolution algorithm."

```{r DE bounds, eval = FALSE}
library(RcppDE)
lower <- rep(-6,length(theta)) %>% setNames(names(theta))
upper <- rep(4,length(theta)) %>% setNames(names(theta))
```


```{r DE control and setup, eval = FALSE}
set.seed(330303)
decontrol <- DEoptim.control(NP=10*length(theta), CR=0.925, F=0.85,
                             itermax=100, storepopfrom=0)
```


```{r Fit with DEoptim, eval = FALSE}
fit2 <- DEoptim(fn=pred, lower=lower,upper=upper, control=decontrol,
                .data=data, yobs=yobs)

data.frame(initial = exp(theta),
           DE = exp(fit2$optim$bestmem),
           newuoa  = exp(fit$par),
           CG = exp(fitt$par)) %>% signif(3)
```

## DA for the plot
```{r DEoptim DA, eval = FALSE}
pops <- lapply(fit2$member$storepop, as.data.frame)
hx <- bind_rows(pops)
hx <- mutate(hx, iteration=rep(1:decontrol$itermax,each=decontrol$NP))
hx <- mutate(hx, pop = rep(1:decontrol$NP, time=decontrol$itermax))
hxm <- gather(hx, variable, value, 1:5) %>% mutate(value = exp(value))
best <- as_data_frame(fit2$member$bestmemit) %>% 
  mutate(iteration = 1:decontrol$itermax)
bestm <- gather(best,variable,value,1:5) %>% mutate(value = exp(value))
```


```{r DEoptim plot, eval = FALSE}
ggplot(data=hxm) + 
  geom_line(aes(iteration,value,group=pop),col="darkslateblue") + 
  geom_line(data=bestm,aes(iteration,value),col="orange",lwd=1) + 
  scale_y_continuous(trans="log", breaks=10^seq(-4,4), name="Parameter value") + 
  facet_wrap(~variable, ncol=2, scales="free_y") 
```






