---
title: "Workbook"
author: "Introduction to `mrgsolve` Workshop"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    css: docs/src/styles/styles.css
params:
  include: FALSE
---

```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', warning=FALSE, message=FALSE)
```

# Setup
```{r}
library(mrgsolve)
library(dplyr)
library(magrittr)
library(tidyr)
options(mrgsolve_mread_quiet=TRUE)


a <- config::get()
```

<hr><div class = spacersm></div>

# Mixture model

* Create a model specification file that simulates a mixture model
* Start with a clearance difference in two populations
    * pop1: 30% with reference CL
    * pop2: 70% with 30% reduction in CL
* Next, try to code a clearance difference in three populations
    * pop1: 10% with reference CL
    * pop2: 70% with 30% reduction in CL
    * pop3: 20% with 50% reduction in CL



```{r, "TRY Mixture model", purl=TRUE, eval=FALSE}

code <- '
$PARAM TVCL = 1, TVV = 20, TVKA = 1.5

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN
double pop = 1;

double CL = TVCL;

if(pop == 2) {
  CL = TVCL;
}

double V = TVV; double KA = TVKA;

'

mod <- mcode("mix", code) %>% update(end=72, delta = 0.1)

mod %>% ev(amt=100) %>% mrgsim(nid = 25) %>% plot
```

Answer for two populations
```{r, "ANSWER Mixture model 2", purl=FALSE, include=params$include}

code <- '

$PLUGIN Rcpp

$PARAM TVCL = 1, TVV = 20, TVKA = 1.5, 
ppop2 = 0.7, CL_pop2 = 0.7

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN
if(NEWIND <= 1) {
  capture pop = 1 + R::rbinom(1, ppop2);
}

double CL = TVCL;

if(pop == 2) {
  CL = TVCL * CL_pop2;
}

double V = TVV; double KA = TVKA;
'

mod <- mcode("mix", code) %>% update(end=72, delta = 0.1)

out <- mod %>% ev(amt=100) %>% mrgsim(nid = 500)

count(as_data_frame(out), pop) %>% mutate(n = n/(sum(n)))
```


Answer for three populations
```{r, "ANSWER Mixture model 3", purl=FALSE, include=params$include}

code <- '

$PLUGIN Rcpp

$PARAM TVCL = 1, TVV = 20, TVKA = 1.5
ppop1 = 0.1, CL_pop2 = 0.7,
ppop2 = 0.7, CL_pop3 = 0.5

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN
if(NEWIND <= 1) {
  double popvar = R::runif(0,1);
  capture pop = 1;
  if(popvar > ppop1) pop = 2;
  if(popvar > (ppop1+ppop2)) pop = 3;
}

double CL = TVCL;

if(pop == 2) {
  CL = TVCL * CL_pop2;
} 
if(pop == 3) {
  CL = TVCL * CL_pop3
}

double V = TVV; double KA = TVKA;
'

mod <- mcode("mix", code) %>% update(end=72, delta = 0.1)

out <- mod %>% ev(amt=100) %>% mrgsim(nid = 500)

count(as_data_frame(out), pop) %>% mutate(n = n/(sum(n)))
```



<hr><div class = spacer></div>
